Slavnov products and KP tau functions

### 2025-04-15 draft.tex
correct integral representation of the tau function

The coefficients \(L_{jk}\) are given by \todo{\tiny Check references
  1, 2 of the paper to see the trigonometric case.}
\begin{equation}
  L_{jk} = g(u_k, \bm{u}_k) \mathcal{Y}(u_k; \bm{u}_j)
\end{equation}
and from this expression one can calculate the matrix \(\bm{M}\).
Finally, the authors also show that if \(\det (\bm{M}) = 0 \) and
\(\mathrm{rank}(\bm{M}) = n\), then, the Slavnov products are given by
\begin{equation}
 \bm{\zeta}_\ell = \phi(\bm{v}) \tilde{\Delta}(\bm{u}_\ell)\hat{\Omega}_\ell\; ,
\end{equation}
where \(\phi(\bm{v})\) is a function of the Bethe roots, and also
\begin{equation}
  \tilde{\Delta}(\bm{u}_\ell) = \prod_{\substack{u_j, u_j \in \bm{u}_\ell \\ j>k}} \frac{c}{u_j - u_k}\; . 
\end{equation}
Therefore, we can absorb the product of the constant \(c\) into
a unique constant \(c_0\) so that 
\begin{equation}
  \frac{\tilde{\Delta}(\bm{u}_\ell)}{c_0} = \prod_{\substack{u_j, u_j \in \bm{u}_\ell \\ j>k}} \frac{1}{u_j - u_k}\; , 
\end{equation}
that is the inverse of the Vandermonde determinant
\(\Delta(\bm{u}_\ell)\).
Finally, we have  that \(\hat{\Omega}_\ell\) are minors of the matrix \(\bm{\Omega}\) with the \(\ell\)-th
column excluded, that is 
\begin{equation}
 \hat{\Omega}_\ell = \det_{k\neq \ell} \Omega_{jk}\; ,
\end{equation}
where
\begin{equation}
  \Omega_{jk} = g(u_k, v_j) \mathcal{Y}(u_k; \{u_k, \bm{v}_j\})
  \qquad j= 1, \dots, n\; , \quad k =1, \dots, n+1
\end{equation}

@@

In fact, we can focus all our attention on a particular function. Therefore,
for a fixed \(\ell\), we define the vector \( \bm{z} \equiv \bm{u}_\ell\).
Consequently, 
\begin{equation}
\begin{split}
\label{eq:tau-function}
  \bm{\tau}(\bm{z}, \bm{v})
  & = \frac{1}{\Delta(\bm{z})} \det[\Omega_j(z_k) ]_{j,k=1}^n \\
  & = \frac{1}{\Delta(\bm{z})} \det\left[\frac{\mathcal{Y}_j(z_k)}{z_k - v_j} \right]_{j,k=1}^n  \\
\end{split}
\end{equation}
All these tau functions have the form of an
\emph{alternant determinant}.

@@

It is easy to show that that the tau function~(\ref{eq:tau-function})
admits an integral representation of the form
\begin{equation}
  \label{eq:tau-function-miwa}
  \bm{\tau}(\bm{t}, \bm{v}) = 
    \det_{j,k}\left(
    \oint_{\gamma_k} \frac{d w}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k} \; ,
    \right)
\end{equation}
where 
\begin{equation}
  \xi(\bm{t}, w^{-1}) = \sum_{p=1}^\infty t_p w^{-p}\; ,
\end{equation}
and the integration contour \(\gamma_k\) encloses all points
\(\{z_j\}_{j=1}^n\) but avoids the Bethe root \(v_k\).  For a detailed
proof of this statement, one may check the
reference~\cite{Araujo:2021ghu}.  The only difference now is that now
we need to be carefull with the integration contour \(\gamma_k\).
One of the advantages of this representation is that now it makes
sense to consider the limit of the infinite chain, \(n\to \infty\).

For later convencience, let us also define an invertible matrix
\(\mathcal{K}\) by its components
\begin{equation}
    \mathcal{K}_{jk} = 
\oint_{\gamma_k} \frac{d w}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k}\;,  
\end{equation}
so that the tau function is simply \(\bm{\tau}(\bm{t}, \bm{v}) = \det\mathcal{K}\).

@@

Using the expression~(\ref{eq:tau-function-miwa}), we can write the
Baker-Akhiezer function in terms of the \(z\)-coordinates.
Let us first write
\begin{subequations}
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t}, \lambda)} & = \exp\left( \sum_{p=1}^\infty t_p \lambda^p \right)
    = \exp\left( \sum_{p=1}^\infty \sum_{j=1}^n \frac{1}{p} z^p \lambda^p \right) \\
    & = \exp\left( \sum_{j=1}^n \sum_{p=1}^\infty  \frac{1}{p} z_j^p \lambda^p \right)
    = \exp\left( - \sum_{j=1}^n \ln( 1 -  z_j \lambda ) \right) \\
    & =  \prod_{j=1}^n \frac{1}{1 -  z_j \lambda} 
  \end{split}
\end{equation}
and we see that these functions have simple poles at \( \lambda = z_j\).

Moreover, 
\begin{equation}
  t_p - \lambda^{-p}/p = \frac{1}{p} \sum_{j=1}^n z_j^ p - \lambda^{-p}/p \; , 
\end{equation}
and we can write
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t} - [\lambda^{-1}], w^{-1})} & = \exp\left[ \sum_{p=1}^\infty \left(t_p  - \frac{\lambda^{-p}}{p}\right) w^{-p} \right]\\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ - \sum_{p=1}^\infty \frac{\lambda^{-p}}{p} w^{-p} \right] \\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ \ln ( 1 - 1 / (\lambda w)) \right]\\
    & = \left( 1 - \frac{1}{\lambda w}\right)  \prod_{j=1}^n \frac{1}{1 -  z_j /w} \; .
  \end{split}
\end{equation}
\end{subequations}

>>>>

The coefficients \(L_{jk}\) are given by \todo{\tiny Check references
  1, 2 of the paper to see the trigonometric case.}
\begin{equation}
  L_{jk} = g(u_k, \bm{u}_k) \mathcal{Y}(u_k; \bm{u}_j)
\end{equation}
and from this expression one can calculate the matrix \(\bm{M}\).
Finally, the authors also show that if \(\det (\bm{M}) = 0 \) and
\(\mathrm{rank}(\bm{M}) = n\), then, the Slavnov products are given by
\begin{equation}
 \bm{\zeta}_\ell = \phi(\bm{v}) \tilde{\Delta}(\bm{u}_\ell)\hat{\Omega}_\ell\; ,
\end{equation}
where \(\phi(\bm{v})\) is a function of the Bethe roots, and also
\begin{equation}
  \tilde{\Delta}(\bm{u}_\ell) = \prod_{\substack{u_j, u_j \in \bm{u}_\ell \\ j>k}} \frac{c}{u_j - u_k}\; . 
\end{equation}
Therefore, we can absorb the product of the constant \(c\) into
a unique constant \(c_0\) so that 
\begin{equation}
  \frac{\tilde{\Delta}(\bm{u}_\ell)}{c_0} = \prod_{\substack{u_j, u_j \in \bm{u}_\ell \\ j>k}} \frac{1}{u_j - u_k}\; , 
\end{equation}
that is the inverse of the Vandermonde determinant
\(\Delta(\bm{u}_\ell) = \prod_{j>k}(u_j - u_k)\).
Finally, we have  that \(\hat{\Omega}_\ell\) are minors of the matrix \(\bm{\Omega}\) with the \(\ell\)-th
column excluded, that is 
\begin{equation}
 \hat{\Omega}_\ell = \det_{k\neq \ell} \Omega_{jk}\; ,
\end{equation}
where
\begin{equation}
  \Omega_{jk} = g(u_k, v_j) \mathcal{Y}(u_k; \{u_k, \bm{v}_j\})
  \qquad j= 1, \dots, n\; , \quad k =1, \dots, n+1
\end{equation}

@@

In fact, we can focus all our attention on a particular function. Therefore,
for a fixed \(\ell\), we define the vector \( \bm{z} \equiv \bm{u}_\ell\).
Consequently, 
\begin{equation}
\begin{split}
\label{eq:tau-function}
  \bm{\tau}(\bm{z}, \bm{v})
  & = \frac{1}{\Delta(\bm{z})} \det[\Omega_j(z_k) ]_{j,k=1}^n \\
  & = \frac{1}{\Delta(\bm{z})} \det\left[\frac{\mathcal{Y}_j(z_k)}{z_k - v_j} \right]_{j,k=1}^n  \; .
\end{split}
\end{equation}
All these tau functions have the form of an \emph{alternant
determinant}.  Henceforth, we will only refer to the matrix
\(\bm{\Omega}\) defined by the components
\begin{equation}
  \label{eq:omega-matrix}
  \Omega_{jk} = \frac{\mathcal{Y}_j(z_k)}{z_k - v_j}\; .
\end{equation}

@@

Let us also define
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t}, \lambda)} & = \exp\left( \sum_{p=1}^\infty t_p \lambda^p \right)
    = \exp\left( \sum_{p=1}^\infty \sum_{j=1}^n \frac{1}{p} z^p \lambda^p \right) \\
    & = \exp\left( \sum_{j=1}^n \sum_{p=1}^\infty  \frac{1}{p} z_j^p \lambda^p \right)
    = \exp\left( - \sum_{j=1}^n \ln( 1 -  z_j \lambda ) \right) \\
    & =  \prod_{j=1}^n \frac{1}{1 -  z_j \lambda} 
  \end{split}
\end{equation}
where \(\lambda\) is a complex parameter. 
and we see that these functions have simple poles at \( \lambda = z_j^{-1}\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Integral representation of the tau functions}

It is easy to show, see~\cite{Araujo:2021ghu} and references therein,
that that the tau function~(\ref{eq:tau-function}) admits an integral
representation of the form
\begin{equation}
  \label{eq:tau-function-miwa}
  \bm{\tau}(\bm{t}, \bm{v}) =
    \det_{j,k}\left(
    \oint_{\gamma_k} \frac{d w}{2\pi i} e^{\xi(\bm{t}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k} \; ,
    \right)
\end{equation}
where 
\begin{equation}
  \xi(\bm{t}, w^{-1}) = \sum_{p=1}^\infty t_p w^{-p} \quad \Leftrightarrow \quad
  \xi(\bm{z}, w^{-1}) = w^n \prod_{j=1}^n (w - z_j)^{-1} \; ,
\end{equation}
and the integration contour \(\gamma_k\) encloses all points
\(\{z_j\}_{j=1}^n\) but avoids the Bethe root \(v_k\).
One of the advantages of this representation is
that now it makes sense to consider the limit of the infinite chain,
\(n\to \infty\).

There are some minor differences between our calculations and the
original references. First, we need to be carefull with the
integration contour \(\gamma_k\). For this reason, we want to
show explicitly this result.

Let us define the invertible matrix \(\bm{\mathcal{K}}^T\) by its components
\begin{equation}
    \mathcal{K}_{kj} = 
\oint_{\gamma_k} \frac{d w}{2\pi i} e^{\xi(\bm{z}, w^{-1})} \frac{w^{-j}\mathcal{Y}_k(w)}{w - v_k}\;. 
\end{equation}
Explicitly, we have 
\begin{equation}
  \begin{split}
    \mathcal{K}_{kj}
    = & \oint_{\gamma_k} \frac{d w}{2\pi i} \frac{w^{n-j}}{\prod_{j} (w - z_j)}
    \frac{\mathcal{Y}_k(w)}{w - v_k} = \sum_{\ell =1}^n \frac{z_l^{n-j}}{\prod_{j\neq \ell} (z_\ell - z_j)}
    \frac{\mathcal{Y}_k(z_\ell)}{z_\ell - v_k}\\
    & = \sum_{\ell = 1}^n  \Omega_{k \ell} \left( \frac{z_\ell^{n-j}}{\prod_{j\neq \ell} (z_\ell - z_j)} \right)
  \end{split}
\end{equation}
where in the second line we have used the
\(\bm{\Omega}\)-matrix defined in~(\ref{eq:omega-matrix}).

It is easy to see that
\begin{equation}
  \begin{split}
  \det \bm{\mathcal{K}} & = \det(\bm{\Omega}) \det_{l,j}\left( \frac{z_l^{n-j}}{\prod_{j\neq l} (z_l - z_j)}\right) \\
  & = \det(\bm{\Omega}) \det_{l, j}(z_l^{n-j}) \prod_{\substack{j, l =1 \\ j \neq l}}^n (z_l - z_j)^{-1} \; .
  \end{split}
\end{equation}
We also have the identities 
\begin{subequations}
  \begin{equation}
	\det_{l,j}(z_l^{n-j}) = (-1)^{n(n-1)/2} \Delta(\bm{z})\; ,
  \end{equation}
  and
  \begin{equation}
	\prod_{\substack{j, l = 1 \\ j \neq l}}^n (z_l - z_j) = (-1)^{n(n-1)/2} \Delta(\bm{z})^2\; . 
  \end{equation}
\end{subequations}

Putting all these facts together, we finally see that 
\(\bm{\tau}(\bm{z}, \bm{v}) = \det\bm{\mathcal{K}}\).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Baker-Akhiezer function}

@@

Using the expression~(\ref{eq:tau-function-miwa}), we can write the
Baker-Akhiezer function in terms of the \(z\)-coordinates.
Moreover, 
\begin{equation}
  t_p - \lambda^{-p}/p = \frac{1}{p} \sum_{j=1}^n z_j^ p - \lambda^{-p}/p \; , 
\end{equation}
and we can write
\begin{equation}
  \begin{split}
    e^{\xi(\bm{t} - [\lambda^{-1}], w^{-1})} & = \exp\left[ \sum_{p=1}^\infty \left(t_p  - \frac{\lambda^{-p}}{p}\right) w^{-p} \right]\\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ - \sum_{p=1}^\infty \frac{\lambda^{-p}}{p} w^{-p} \right] \\
    & = e^{\xi(\bm{t}, w^{-1})} \exp\left[ \ln ( 1 - 1 / (\lambda w)) \right]\\
    & = \left( 1 - \frac{1}{\lambda w}\right)  \prod_{j=1}^n \frac{1}{1 -  z_j /w} \; .
  \end{split}
\end{equation}
